### 2012-01-01 manuscript.tex
Reflow text

% Template for PLoS
% Version 2.0 July 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
%
% -- FIGURES AND TABLES
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text. For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
%
% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
% For bolding characters in mathmode, please use $\mathbf{}$
%
% Please add line breaks to long equations when possible in order to fit our 2-column layout.
%
% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
%
% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
%
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{graphicx,amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}

%% Include all macros below. Please limit the use of macros.

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{????}
}
% Insert Author names, affiliations and corresponding author email.
\\
Mason Liang$^{1,\ast}$,
Rasmus Nielsen$^{2}$
\\
\bf{1} Mason Liang Integrative Biology, University of California, Berkeley, California, United States of America
\\
\bf{1} Rasmus Nielsen Integrative Biology, University of California, Berkeley, California, United States of America
\\
$\ast$ E-mail: wmliang@berkeley.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
Estimating admixture histories is crucial for understanding the genetic diversity we see in present-day populations. Existing allele frequency or phylogeny-based methods are excellent for inferring the existence of admixture or its proportions, but have less power for estimating admixture times. Recently introduced approaches for estimating these times use spatial information from admixed chromosomes, such as the local ancestry or the decay of admixture linkage disequilibrium (ALD).  One popular method, implemented in the programs ALDER and ROLOFF, uses two-locus ALD to infer the time of a single admixture event, but is only able to estimate the time of the most recent admixture event based on this summary statistic.  We derive analytical expressions for the expected ALD in a three-locus system and provide a new statistical method based on these results that is able to resolve more complicated admixture histories. Using simulations, we show how this new statistic behaves on a range of admixture histories. As an example, we also apply our method to the Colombian and Mexican samples from the 1000 Genomes project.

\section*{Introduction}
There are many methods for inferring the presence of admixture, e.g. methods
using simple summary statistics detecting deviations from phylogenetic symmetry
\cite{reich2009reconstructing} [TODO(mason) Also cite patterson paper and paper
from Monty's group], methods estimating admixture proportions with programs such
as Structure \cite{pritchard2000inference}, Admixture \cite{alexander2009fast}
or RFmic [cite] [TODO comment: we might want to make sure to add more citations
here, for example all the ones from our previous papers and the ones from the
ALDER paper. People like to see themselves cited]. However, there has been less
research on estimating admixture times, possibly because such methods require
data which were unavailable until the advent of high-throughput next generation
sequencing. Some recently developed methods use the inferred local ancestry of
sequences to construct admixture tract length distributions [TODO cite Pool Nielsen,
Gravel and yourself]. Over time, recombination is expected to decrease the
average lengths of admixture tracts. The length distribution of admixture tracts
is therefore informative about the time since admixture.  Much of the theory
relating to tracts lengths is based on Fisher's famous theory of junctions
\cite{fisher1949theory} and subsequent work by Stam \cite{stam1980distribution}
and others [TODO add as many citations here as possible]. For example,
\cite{baird2003distribution} first discussed the length distribution of tracts
descended from a single ancestor. These results informed later analyses of
admixture tract length distribution, such as \cite{pool2009inference},
\cite{gravel2012population}, and \cite{liang2014lengths}.
\cite{gravel2012population} also implemented the software program TRACTS, which
estimates admixture histories by fitting the tract length distribution, obtained
by local ancestry inference, to a exponential approximation.

Another approach, which we will follow in this paper, is based on the decay of
admixture linkage disequilibrium (ALD). In a well-mixed, genetically isolated
human populations, linkage disequilibrium decays to zero on a scale of tenths of
centiMorgans. However, when an admixed population is founded, it begins with
large of amount of linkage disequilibrium, which is a result of the allele
frequency differences between the source populations. This occurs even if the LD
in the source populations themselves is negligible. The linkage disequilibrium
in the admixed population then fluctuates in the generations after its founding,
decreasing as a result of drift and recombination, or increasing because of
additional waves of migration. From the LD present in a modern day admixed
population, it is possible to make inferences about the population's admixture
history. This insight was first used in the program ROLLOFF
\cite{moorjani2011history} and was later extended by ALDER
\cite{loh2013inferring}.

These two methods use the fact that if an admixed population takes in no
additional migrants after the founding generation, the LD present in the
population is expected to decay exponentially as a function of distance. The
rate constant of this exponential decay is proportional to the age of the
founding admixture pulse and so can be used as an estimator. ROLLOFF and ALDER
are well suited for inferring the time of the admixture event when the
population's admixture history can be approximated as a single pulse. However,
it can be important to estimate parameters for admixture histories involving
multiple pulses, such as estimating the date of Native American admixture in
Rapa Nui \cite{moreno2014genome} or determining migration patterns in the
Americas \cite{gravel2013reconstructing}. In these instances the expected decay
of LD will become a mixture of exponentials. ROLLOFF and ALDER have limited
resolution, as they can usually only infer the date of the most recent migration
wave \cite{moorjani2011history}, or reject the hypothesis of a single pulse
admixture \cite{loh2013inferring}.

ROLLOFF and ALDER use the information contained in pairs of sites by examining
the two-locus linkage disequilibrium between them. Here we extend the theory
underlying the methods in ROLOFF and ADLER to three loci by considering
three-locus LD. There are two ways of measuring the linkage between $n$ loci.
\cite{bennett1952theory} defines $n$-locus linkage in a way that maintains a
geometric decrease of LD each generation as a result of recombination, which is
an important property of two-locus linkage disequilibrium.
\cite{slatkin1972treating} defines $n$-locus LD  to be the $n$-way covariance,
analogously to the property of two locus LD as the covariance in allele
frequency between pairs of loci. For two and three loci, these two definitions
coincide, but for four or more loci, they do not.

In this paper, we will use Bennett and Slatkin's definition of three-locus LD to
examine the the decay of ALD for three sites as a function of the genetic
distance between them. We derive an equation that describes the decay of
three-locus LD under an admixture history with multiple waves of migration. We
then compare the results of coalescent simulations to this equation, and develop
some guidelines for when admixture histories more complex than a single pulse
can be resolved. Finally, we apply our method to the Colombian and Mexican
samples in the 1000 Genomes data set, using the Yoruba samples as a reference.
Fitting a two-pulse model to data, we estimate admixture histories for the two
populations which are qualitatively consistent with the results reported in
\cite{gravel2013reconstructing}.

[TODO Could we say anything here about Neanderthals in Asia?]

% You may title this section "Methods" or "Models".
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis"
\section{Model}
We use a random union of gametes admixture model as described in
\cite{liang2014understanding}, which is an extension of the mechanistic
admixture model formulated by \cite{verdu2011general}. In this model, two or
more source populations contribute migrants to form an admixed population
consisting of $2N$ haploid individuals. Each generation in the admixed
population is formed through the recombination of randomly selected individuals
from the previous generation, with some individuals potentially replaced by
migrants from the source populations. For simplicity, we consider a model with
only two source populations. Furthermore, the first source population only
contributes migrants in the founding generation, $T$. The second source
population contributes migrants in the founding generation and possibly in one
or more generations thereafter. In generation $i$, for $i=T-1,\dots,0$ (before
the present), a fraction $m_i$ of the admixed population is replaced by
individuals from the second source population.

\section{Linkage Disequilibrium and Local Ancestry}
ROLLOFF and ALDER use the standard two-locus measure of LD between a SNP at
positions $x$ and another SNP at position $y$, which is a genetic distance $d$
to the right,

\begin{align}
	D_2(d) = \text{cov}(H_x,H_y),
	\label{D2}
\end{align}

where $H_x$ and $H_y$ represent the haplotype or genotypes of an admixed
chromosome at positions $x$ and $y$. In the case of haplotype data, $H_{i,x}=1$
if the $i^\text{th}$ sample is carrying the derived allele at the SNP at
position $x$, and is otherwise 0. Alternatively, for genotype data, $H_{i,x}$
take on values from $\{0,1/2,1\}$ depending on the number of copies of the
derived allele the $i^\text{th}$ sample is carrying at SNP position $x$.
We consider an additional site at position $z$, which is located a further
genetic distance $d'$ to the right of $y$. The three-loci LD, as defined by as
defined by \cite{bennett1952theory} and \cite{slatkin1972treating}, is given by

\begin{align}
	D_3(d,d')
	= \text{cov}(H_x,H_y,H_z)
	= \mathbb{E}[(H_x-\mathbb{E}H_x)(H_y-\mathbb{E}H_y)(H_z-\mathbb{E}H_z)].
\label{D3}
\end{align}

The LD in an admixed population depends on the genetic differentiation between the source populations and and its admixture history. Let $A_x$ represent the local ancestry at position $x$, with $A_x=1$ if $x$ is inherited from an ancestor in the first source population, and $A_x=0$ if $x$ is inherited from the second source population. We can compute $D_3$ in terms of the three-point covariance function of $A_x$ and so separate out the effects of allele frequencies and local ancestry. Let $H_x = f_x+\delta A_x$, where $f_x$ is the allele frequency of locus $x$ in the first source population and $\delta_x$ is the difference of the allele frequencies of locus $x$ in the two source populations. We now make the assumption that the allele frequencies in the source populations are known and fixed. Equation \ref{D3} then becomes

\begin{align}
	D_3(d,d')
		&= \text{cov}\left(f_x+\delta_xA_x,f_y+\delta_yA_y,f_z+\delta_zA_z\right)\nonumber \\
		&= \delta_x\delta_y\delta_z \text{cov}(A_x,A_y,A_z).
	\label{local_ancestry}
\end{align}
A similar argument shows that $D_2(d)$ is proportional to the two-point covariance function of the local ancestry.

\subsection{Local Ancestry Covariance Functions}
From the above section we see that we can describe the three-point admixture LD in terms of covariances of local ancestry in the three points.  We now expand the covariance in equation \ref{D3} into its component expectations to get
\begin{align*}
	\text{cov}(A_x,A_y,A_z)
		&= \mathbb{E}[A_xA_yA_z]
		-\mathbb{E}[A_xA_y]\mathbb{E}[A_z]
		-\mathbb{E}[A_xA_z]\mathbb{E}[A_y]
		-\mathbb{E}[A_yA_z]\mathbb{E}[A_x]
		+2\mathbb{E}[A_x]\mathbb{E}[A_y]\mathbb{E}[A_z].
\end{align*}
Each one of these expectations on the right-hand side is the probability that one or more sites is inherited from an ancestor from first source population. We organize these products of probabilities in a column vector:
\begin{align*}
	\mathbf{v}_3 &= \left(\begin{array}{l}
	\mathbb{P}\{A_x=A_y=A_z=1\}\\
	\mathbb{P}\{A_y=A_z=0\}\mathbb{P}\{A_x=0\}\\
	\mathbb{P}\{A_x=A_z=0\}\mathbb{P}\{A_y=0\}\\
	\mathbb{P}\{A_x=A_y=0\}\mathbb{P}\{A_z=0\}\\
	\mathbb{P}\{A_x=0\}\mathbb{P}\{A_y=0\}\mathbb{P}\{A_z=0\}
	\end{array}\right),
\end{align*}
so that $\text{cov}(A_x,A_y,A_z) = (1,-1,-1,-1,2)\mathbf{v}_3$. There is one entry in $\mathbf{v}_3$ for each of the five ways in which the three markers at positions $x,y$, and $z$ can arranged on one or more chromosomes. In the founding generation $T$, this column vector is given by $\mathbf{v}_{3(T)} = (1-m_T,(1-m_T)^2,(1-m_T)^2,(1-m_T)^2,(1-m_T)^3)'$. The probabilities for subsequent generations can be found by left-multiplying drift, recombination, and migration matrices:
\begin{align*}
	\mathbf{v}_{3(i)} = \mathbf{D}_i \mathbf{L}\mathbf{U} \mathbf{v}_{3(i-1)},
\end{align*}
The matrices $\mathbf{D}_i$, $\mathbf{L}$, and $\mathbf{U}$ account for the effects of migration, drift, and recombination, respectively. The migration matrix is a diagonal matrix given by
\begin{align*}
	\mathbf{D}_i &= \text{diag}(1-m_i,(1-m_i)^2,(1-m_i)^2,(1-m_i)^2,(1-m_i)^3).
\end{align*}
Its entries are the probabilities that one, two, or three chromosomes in the admixed population will not be replaced by chromosomes from the second source population in generation $i$. The lower triangular drift matrix
\begin{align*}
	\mathbf{L}&=\frac{1}{4N^2}\left( \begin{array}{ccccc}
					 4N^2 		& 0 	& 0 & 0 & 0\\
					 2N 	& 2N(2N-1) & 0 & 0 & 0\\
					 2N 	& 0 & 2N(2N-1) & 0 & 0\\
					 2N 	& 0 & 0 & 2N(2N-1) & 0\\
					 1 & 2N-1 & 2N-1 & 2N-1 & (2N-1)(2N-2)
				\end{array} \right)
\end{align*}
gives the standard Wright-Fisher drift transition probabilities between the states as a function of the population size $2N$. Finally, the upper triangular recombination matrix is determined by the recombination rates between the three sites:
\begin{align*}
	\mathbf{U} &=
		\left( \begin{array}{ccccc}
		e^{-d-d'} & (1-e^{-d})e^{-d'} & (1-e^{-d})(1-e^{-d'}) & e^{-d}(1-e^{-d'}) & 0 \\
		0 & e^{-d'} & 0 & 0 & 1-e^{-d'} \\
		0 & 0 & 1-e^{-d}-e^{-d'}+2e^{-d-d'} & 0 & e^{-d}+e^{-d'}-2e^{-d-d'} \\
		0 & 0 & 0 & e^{-d} & 1-e^{-d} \\
		0 & 0 & 0 & 0 & 1 \end{array} \right)
\end{align*}
The covariance function is then given by
\begin{align}
	\text{cov}(A_x,A_y,A_z) &= \left(1,-1,-1,-1,2\right)\left(\prod_{i=0}^{T-1} \textbf{D}_i \mathbf{L}\mathbf{U}\right)\textbf{v}_{3(0)}.
	\label{cov}
\end{align}
We can obtain an analogous equation for $\text{cov}(A_x,A_y)$, involving the migration, drift, and recombination matrices for two loci:
$$
	\text{cov}(A_x,A_y) = \left(1,-1\right)\left(\prod_{i=0}^{T-1} \textbf{D}_i \mathbf{L}\mathbf{U}\right)\textbf{v}_{2(0)}.
$$

In some cases, equation \ref{cov} simplifies further. In a one-pulse migration model, in which $m_T=M$ and is there after 0, the $\mathbf{D}_i$'s become identity matrices, and we get the closed from expression
$$
	\text{cov}(A_x,A_y,A_z) = M(1-M)(1-2M)\left(1-\frac{1}{2N}\right)^T\left(1-\frac{2}{2N}\right)^Te^{-T(d+d')}.
$$
This is because $(1,-1,-1,-1,2)$ is a left eigenvector of both $\mathbf{L}$ and $\mathbf{U}$, with corresponding eigenvalues $(1-1/2N)(1-2/2N)$ and $\exp(-d-d')$. Note that when $M=0$, the covariance function will be identically 0. Another case is a two pulse model in which we ignore the effects of genetic drift. In this model, admixture only occurs $T$ and $T_2$ generations before the present, so that $m_T=M_1,m_{T'}=M_2$, and all other $m_i$'s are 0. Making the substitution $T_1=T-T_2$, the right hand side of equation \ref{cov} becomes
\begin{align}
		&(1-M_1)(1-M_2)e^{-T_2(d+d')}\left[M_2(1-M_1)^2-2M_2^2(1-M_1)^2+M_1(1-2M_1)e^{-T_1(d+d')}\right.\nonumber\\
		&\ \ \left.-M_1M_2(1-M_1)\left(e^{-M_1d}+e^{-M_1 d'}+\left(1-e^{-d}-e^{-d'}+2e^{-d-d'}\right)^{T_1}\right) \right].
\label {ash_2p}
\end{align}
The corresponding expression for the two-point covariance function is given by
\begin{align}
		(1-M_1)(1-M_2)e^{-T_2d}\left(M_2-M_1M_2+M_1e^{-T_1d}\right),
\label{alder_2p}
\end{align}
which is a mixture of two exponentials.

\section{Weighted Linkage Disequilibrium}
As \cite{loh2013inferring} noted, we cannot use the LD in the admixed population directly, because the allele frequency differences in the source populations can be of either sign. As in As \cite{loh2013inferring} , we solve this problem by computing the product of the values of the three-point linkage disequilibrium coefficient with the product of the allele frequency differences. Using equation \ref{local_ancestry} we obtain
$$
\delta_x\delta_y\delta_z D_3(d,d')=\delta_x^2\delta_y^2\delta_z^2\mathbb{E}[\text{cov}(A_x,A_y,A_z)],
$$
because the local ancestry in the admixed sample is independent of the allele frequencies in the admixed population. For inference purposes, we estimate this function by averaging over triples of SNPs which are separated by distances of approximately $d$ and $d'$. The LD term is estimated from the admixed population, while the $\delta$'s are estimated from reference populations which are closely related to the two source populations. We notice that both this approach, as well as the previous approaches (e.g., \cite{loh2013inferring} ), does not take genetic drift in the source populations after the time of admixture into account, i.e. there is an assumption of both this mehtod and previous methods that the allele frequencies in the ancestral source populations can be approximated well using the allele frequencies in the extant populations.

We arrange the data from the admixed samples in an $n\times S_n $ matrix $\mathbf{H}$, where $n$ is the number of admixed haplotypes/genotypes, and $S_n$ is the number of segregating sites in the sample. For ease of notation, we assume that the positions are given in units which make the unit interval equal to the desired bin resolution.

For a given $d$ and $d'$ the set of SNP triples we use in the estimator for the weighted LD is
$$
	S[d,d'] = \left\{x,y,z: d\leq  x-y < d+1 \text{ and } d' \leq y-z < d'+1 \right\}.
$$
Let $f_x$ be empirical allele frequency in the admixed population. An unbiased estimator of the weighted three-point linkage disequilibrium coefficient is then
\begin{align*}
	\hat{a}[d,d']&=
		\frac{1}{|S[d,d']|}
		\sum_{x,y,z\in S[d,d']}\frac{n\sum_{i=1}^n
			\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)}{(n-1)(n-2)}.
\end{align*}

\section{Algorithm}
Directly computing $\hat{a}[d,d']$ over the set $d,d'\in \{0,1,\dots,P\}^2$ would be cubic in the number of segregating sites, but as is the case with ALDER, we can use using a fast Fourier transform (FFT) to approximate $\hat{a}$, giving an algorithm whose run-time is instead linear in the number of segregating sites. We first rearrange $\hat{a}$ to get
\begin{align*}
	\hat{a}[d,d']
		&=\frac{n}{(n-1)(n-2)}
		\frac{\sum_{i=1}^n\sum_{x,y,z\in S[d,d']}
			\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)}
		{\sum_{x,y,z\in S[d,d']}1},
\end{align*}
and define sequences $b_i[d]$ and $c[d]$ by binning the data and then doubling the length by padding with $P$ zeros,
\begin{align*}
	b_i[d] &= \left\{
     \begin{array}{ll}
       \sum_{x: d\leq \lfloor{x}\rfloor < d+1}\delta_x(H_{i,x}-f_x) & : 0\leq d < P\\
       0 & : P\leq d <2P
     \end{array}
   \right.\\
  	c[d] &= \left\{
     \begin{array}{ll}
       |\{x: d\leq \lfloor{x}\rfloor < d+1\}| & : 0\leq d < P\\
       0 & : P\leq d <2P
     \end{array}
   \right.
\end{align*}
We can approximate $|S[d,d']|$ and the $n$ sums in the numerator of $\hat{a}[d,d']$ in terms of convolutions of these sequences:
$$
	|S[d,d']| \approx \sum_{w=0}^P c[w]c[w+d]c[w+d+d']
$$
$$
	\sum_{x,y,z\in S[d,d']}\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)
		\approx
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'].
$$
These convolutions can be efficiently computed with an FFT, since under a two-dimensional discrete Fourier transform,
$$
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'] \leftrightarrow B_i[j]\bar{B_i}[k]B_i[k-j],
$$
where $B_i$ is the (one-dimensional) discrete Fourier transform of $b$, and $B_i[-j]$ is the $j^\text{th}$ to last most element of $B_i$. Summing over $i$ and taking the inverse discrete Fourier transform, we can approximate the discrete Fourier transform of the numerator of  $\hat{a}$. We use the same method applied to $c$ to approximate the denominator of $\hat{a}$. Because the number of bins is generally much less than the number of segregating sites, the rate-limiting step of this algorithm is the binning step to form $c$ and the $b_i$'s, which is $O(S_n)$, rather than the FFTs, which are $O(P^2\log(P))$.

When using only the admixed population itself as a reference population, the method described above will be biased if the same samples are used to estimate both the linkage disequilibrium coefficients and the weights ($\delta_x$, $\delta_y$, and $\delta_z$). We cannot efficiently compute a polyache statistics like \cite{loh2013inferring}. At the cost of some power, we instead adopt the approach of \cite{pickrell2012inference} and separate the admixed population into two equal-sized groups. We then use one group to estimate the weights, and the other group to estimate linkage disequilibrium coefficients, and vice versa. This gives gives two unbiased estimates for the numerator of $\hat{a}$, which we then average. [TODO Comment: I didn't understand how this works.  Coudl we five more detail?]
\subsection{Fitting the Two-Pulse Model}
We fit equation \ref{alder_2p} to the estimates of the weighted LD using non-linear least squares, with two modifications. We added a proportionality constant to account for the expected square allele frequency difference between the source populations. We also subtracted out an affine term in the weighted LD which is due to population substructure \cite{loh2013inferring}. We estimated this by computing the three-way covariance between triples of chromosomes. We use the jackknife to obtain confidence intervals for the resulting estimates by leaving out each chromosome in turn and refitting on the data for the remaining chromosomes.  [TODO Comment: again, we need more detail here. These are important parts of the manuscript]
\section{Simulations}
We used the program macs \cite{chen2009fast} to generate two source populations which diverged 4000 generations ago and a coalescent simulation to generate an admixed population from the two source populations according to two-pulse and constant admixture models. We sampled 50 diploid individuals from the admixed and two source populations, each consisting of 20 chromosomes of length 1 Morgan. The effective population size was $2N=1000$ for the admixed population and two source populations. Using a two pulse model, we varied the migration probabilities and timings for each pulse to examine the accuracy of equation \ref{alder_2p}. We also simulated data for a model with a constant rate of admixture each generation, and compared this to the predictions made by equation \ref{cov}.
\section{Data Set}
We computed the weighted LD for the Mexican and Columbia populations in the first phase of the 1000 Genomes data set. These consisted of 66 individuals from Los Angeles and 60 individuals from Medellin, respectively. We used the 88 Yoruba samples as the one reference population. We computed the weighted LD on the genotypes to avoid effects of phasing errors.
% Results and Discussion can be combined.
\section{Results and Discussion}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\subsection{Simulations}

\begin{figure}
\includegraphics[scale=.6]{Ts.eps}\caption{
{\bf Predicted weighted LD surfaces from simulations and theory for varying admixture times.} The heat maps are from simulations and the contours are plotted from equation \ref{D3}. The two admixture probabilities were fixed at $m_1=m_2=.2$ and the the times of the two admixture pulses, $T_1$ and $T_2$, were varied. Each square covers the range $0.5 \text{ cM }<d,d'<20\text{ cM}$. When time of the more recent pulse is greater than half of that of the more ancient pulse, i.e. $2T_1>T_1+T_2$, the contours of the resulting weighted LD surface are straight, making it difficult to distinguish from the weighted LD surface produced by a one-pulse admixture scenario.
}
\label{Ts}
\end{figure}

[TODO: the color coding should be labeled for all figures.  Also, the y and x-axes should be labeled]

\begin{figure}
\includegraphics[scale=.6]{ms.eps}
\caption{
{\bf Predicted weighted LD surfaces from simulations and theory for varying admixture proportions.} The heat maps are from simulations and the contours are plotted from equation \ref{D3}. The two admixture times were fixed at 2 and 12 generations ago ($T_1=10$ and $T_2=2$) while the admixture probabilities were varied. Each square covers the range $0.5 \text{ cM }<d,d'<20\text{ cM}$. As the total admixture proportion $m_2+m_1(1-m_2)$ increases above 0.5, the contours change to reflecting that the majority contribution of the genetic material now originates from the other population. Weighted LD surfaces for $m_1>0.5$ or $m_2>0.5$ are not shown, but are qualitatively similar to the surfaces on the lower and rightmost sides.
}
\label{ms}
\end{figure}

\begin{figure}
\includegraphics[scale=.8]{continuous.eps}
\caption{
{\bf Weighted LD surfaces produced by constant admixture.} The heat maps are from simulations and the contours are from equation XX. In all six plots, admixture stopped 5 generations before the present. Each square covers the range $0.5 \text{ cM }<d,d'<20\text{ cM}$. We varied the time of the beginning of the admixture and the total admixture probability. The admixture probability for each generation was constant, and chosen so that the total admixture proportion was either $0.3$ or $0.7$. When the admixture is spread over 5 generations (the leftmost column), the resulting weighted LD surface is similar to a one-pulse weighted LD surface. For longer durations, the weighted LD surfaces are similar to those produced by two pulses of admixture.
}
\label{continuous}
\end{figure}

We first evaluate the accuracy of the equations developed in this paper by comparing the analytical results to simulated data (Figures 1-3).  We find there is a generally a close match between our equations and the simulated data under both the two-pulse admixture scenarios (figures \ref{Ts} and \ref{ms}) and constant-admixture scenarios (figure \ref{continuous}). The exception is when the total admixture proportion $M_2+M_1(1-M_2)$ is close to 0.5. As the total admixture proportion increases above 0.5, the contours for equation \ref{D3} flip from being concave down to concave up. This transition can be seen by comparing the upper left side of figure \ref{ms} to its lower right. At this threshold, the contours of the estimated weighted LD depend on the actual admixture fractions of the samples, which may differ from the expectation as a result of genetic drift. This mismatch between theory and simulations is most evident in figure \ref{ms}, for $m_1=0.1,m_2=0.4$ and $m_1=0.2,m_2=0.4$.

When there is continuous admixture scenario, the shape of the weighted LD surface depends on both the duration and total amount of admixture. When the duration is short, the weighted LD surfaces are indistinguishable from the weighted LD surfaces produced by one pulse of migration. As the duration increases, the contours of the weighted LD surface become more curved. The contours are concave up when the total proportion is greater than $50\%$ and concave down when it is less. When the total proportion is exactly $50\%$, the amplitude of the weighted LD surface is much smaller than the sampling error.

For two pulse models, the effects of the second pulse of migration only become evident when temporal spacing between the pulses is large enough ($T1 > T2$). Otherwise, the resulting weighted LD surface cannot be distinguished from the weighted LD surface produced by one pulse of admixture. As in the case of continuous admixture the concavity of the surface contours is determined by the total admixture proportion.

\begin{figure}
\includegraphics[scale=.6]{estimates.eps}
\caption{
{\bf Accuracy of estimates of $T_1$ as a function of other parameters.} Nine admixture scenarios, $T_1\in \{5,10,20\}$ and $T_2\in\{2,5,10\}$, were simulated 100 times each. The admixture probabilities were fixed at $M_1=0.3$ and $M_2=0.2$. The colored bars give the medians of estimates for each of these nine cases, the boxes delimit the interquartile range, and the whiskers extend out to 1.5 times the interquartile range. As the time between the two pulses of admixture increases, the error in the estimates decreases. Consistent with the simulations shown in figure \ref{Ts}, there is limited power to estimate the time of the more ancient admixture pulse when $T_2>T_1$.
}
\label{estimates}
\end{figure}

[TODO Comment: we need  figures showing the similar one-dimensional LD for comparison.  Maybe the first thing to discuss is to show how you can get approximately the same 1D curve for different models which produce different 2D surfaces]

We next evaluate the utility of the method for estimating admixture times.  The qualitative similarities between one pulse and two pulse admixture scenarios seen in the previous  simulations under some parameter settings, will naturally affect  the estimates. As shown in figure \ref{estimates}, when the spacing between the two pulses is small relative to their age, the median of the estimates of the timing of the second pulse is close to the true value, but the interquartile range is large. Moreover, the best fit often lies on a boundary of the parameter space which is equivalent to a one pulse admixture model. When the spacing between the pulses is larger, the estimates for the timing of the older pulse becomes more precise.
\subsection{1000 Genomes}
\begin{figure}
\includegraphics[scale=.6]{MXL.eps}
\caption{
{\bf Weighted LD surface for Mexican samples with Yoruba as reference.} The model with the best fit is two pulses from the non-Yoruba source population at $T_1+T_2=12.3 \pm 3.3$ and $T_2=9.9\pm 2.7$ generations ago. The jackknife confidence intervals for the times of these two pulses overlap.
}
\label{ASH_MXL}
\end{figure}

To illustrate the utility of the method we we computed weighted LD surfaces for Mexican and Columbian samples from the 1000 Genomes consortium previously analyzed for similar purposes by \cite{gravel2013reconstructing}. For the Mexican samples, \cite{gravel2013reconstructing} found a small but consistent amount of African ancestry, which appeared in the population 15 generations ago, with continuing contributions from European and Native American populations since that date, but no African migration. In fitting a two-pulse model to the Mexican weighted LD surface (figure \ref{ASH_MXL}), we estimated that the two pulses occurred $12.3\pm3.3$ and $9.9\pm2.7$ generations ago. These confidence intervals overlap, and so we cannot reject a one-pulse admixture history. This is not quite consistent with the constant migration model that \cite{gravel2013reconstructing} found, but as we have seen from simulations, it is hard to distinguish a constant migration model from a one-pulse model when the duration of the migration is short.

\begin{figure}
\includegraphics[scale=.6]{CLM.eps}
\caption{
{\bf Weighted LD surface for Colombian samples with Yoruba as reference} The two-pulse model that fits best is two pulses of non-Yoruba admixture at $T_1+T_2=11.8\pm 1.2$ and $T_2=2.64 \pm 0.08$ generations ago. The jackknife confidence intervals for the times of these two pulses do not overlap. The amplitude of this weighted LD surface is approximately ten times larger than that of the Mexican samples. This a result of larger proportion of Yoruba ancestry in the Colombian samples.
}
\label{ASH_CLM}
\end{figure}

The weighted LD surface for the Columbia samples is shown in figure \ref{ASH_CLM}. From this, we estimated two pulses of non-Yoruba migration at $11.8\pm 1.2$ and $2.64\pm0.08$ generations before the present. \cite{gravel2013reconstructing} also inferred two pulses of admixture, corresponding to 3 and 9 generations ago. The weighted LD surface of the Colombian samples has contours which are strongly concave up, in contrast to those of the Mexican samples.
\subsection{Comparison to Existing Methods}
Compared to existing weighted LD methods, our our method uses more information in the data because it compares triples of SNPs instead of pairs. This gives our method the ability to infer admixture histories more complex than a one-pulse model. However, this comes at the price of greater estimation variances. ALDER and ROLLOFF can make estimates from just tens of samples, while our method requires hundreds of samples. Part of this difference can be attributed to the fact that ALDER and ROLLOFF make inferences over a smaller class of models, but the main reason arises from the fact that the existing two methods are estimating second moments of the data, while we are estimating third moments. The variance of these estimates are both inversely proportional to the sample size, but the constants for estimating third moments are larger. As data becomes more readily available, this disadvantage should disappear.

\bibliography{thesis}

\begin{figure}
\includegraphics[scale=.6]{alder_Ts.eps}
\caption{
{\bf Extra figure} Corresponding alder curves for two-pulse admixture with varying pulse times. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Red lines are $T_2=2$, Green lines $T_2=5$, and blue lines are $T_2=10$.
}
\end{figure}

\begin{figure}
\includegraphics[scale=.6]{alder_ms.eps}
\caption{
{\bf Extra figure} Corresponding alder curves for two-pulse admixture with varying migration probabilities. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Increasing red values are higher $M_1$'s and increasing green values are higher $M_2$'s  (I will probably need to choose a better color scheme for this).
}
\end{figure}

\end{document}