### 2015-08-25 manuscript.tex
First commit

% Template for PLoS
% Version 2.0 July 2014
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
%
% -- FIGURES AND TABLES
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For sideways tables, use the {rotating} package and use \begin{sidewaystable} instead of \begin{table} in the appropriate section. PLOS guidelines do not accomodate sideways figures.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS	
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment, and for any superscripts or subscripts also include the base number/text. For example, use $mathrm{mm}^2$ instead of mm$^2$ (do not use \textsuperscript command).
%
% DO NOT USE the \rm command to render mathmode characters in roman font, instead use $\mathrm{}$
% For bolding characters in mathmode, please use $\mathbf{}$ 
%
% Please add line breaks to long equations when possible in order to fit our 2-column layout. 
%
% For inline equations, please do not include punctuation within the math environment unless this is part of the equation.
%
% For spaces within the math environment please use the \; or \: commands, even within \text{} (do not use smaller spacing as this does not convert well).
%
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{graphicx,amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}

%% Include all macros below. Please limit the use of macros.

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{????}
}
% Insert Author names, affiliations and corresponding author email.
\\
Mason Liang$^{1,\ast}$, 
Rasmus Nielsen$^{2}$
\\
\bf{1} Mason Liang Integrative Biology, University of California, Berkeley, California, United States of America
\\
\bf{1} Rasmus Nielsen Integrative Biology, University of California, Berkeley, California, United States of America
\\
$\ast$ E-mail: wmliang@berkeley.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Introduction}
Estimating admixture histories is crucial for understanding the genetic diversity we see in present-day populations. Existing allele frequency or phylogeny-based methods are excellent for inferring the existence of admixture or its proportions, but have less power for estimating admixture times. Recently introduced approaches for estimating these times use spatial information from admixed chromosomes, such as the local ancestry or the decay of linkage disequilibrium (LD). We introduce such a method, based on weighted three-locus LD. In comparison to existing methods based on weighted two-locus LD, our new method is able to resolve more complicated admixture histories. Using simulations, we show how this new statistic behaves on a range of admixture histories. We also apply our method to the Colombian and Mexican samples from the 1000 Genomes project and infer admixture histories consistent with those previously reported.
There are many methods for inferring the presence of admixture, e.g. with $f$-statistics \cite{reich2009reconstructing} or by estimating admixture proportions with programs such as Structure \cite{pritchard2000inference} or Admixture \cite{alexander2009fast}. However, there has been less research on estimating admixture times, possibly because such methods require data which was unavailable until the advent of high-throughput next generation sequencing. Some of these methods use the inferred local ancestry of sequences to construct admixture tract length distributions. Over time, recombination is expected to decrease the average lengths of admixture tracts. This was tract length distribution first worked out in the context of junctions \cite{fisher1949theory} and later extended to randomly mating populations by \cite{stam1980distribution}. \cite{baird2003distribution} first discussed the lengths of tracts descended from a single ancestor. These results informed later analyses of admixture tract length distribution, such as \cite{pool2009inference}, \cite{gravel2012population}, and \cite{liang2014lengths}. \cite{gravel2012population} also implemented the software program TRACTS, which estimates admixture histories by fitting the tract length distribution, obtained by local ancestry inference, to a exponential approximation.

Another approach, which we will follow in this paper, is based on the decay of ancestral linkage disequilibrium (LD). In a well-mixed, genetically isolated human populations, linkage disequilibrium decays to zero on a scale of tenths of centiMorgans. However, when an admixed population is founded, it begins with large of amount of linkage disequilibrium, which is a result of the allele frequency differences between the source populations. This occurs even if the LD in the source populations themselves is negligible. The linkage disequilibrium in the admixed population then fluctuates in the generations after its founding, decreasing as a result of drift and recombination, or increasing because of additional waves of migration. From the LD present in a modern day admixed population, it is possible to make inferences about the population's admixture history. This technique was first in the program ROLLOFF \cite{moorjani2011history} and was later extended by ALDER \cite{loh2013inferring}.

These two methods use the fact that if an admixed population takes in no additional migrants after the founding generation, the LD present in the population is expected to decay exponentially as a function of distance. The rate constant of this exponential decay is proportional to the age of the founding admixture pulse and so can be used as an estimator. ROLLOFF and ALDER are well suited for inferring the time of the admixture event when the population's admixture history can be approximated as a single pulse. However, it can be important to estimate parameters for admixture histories involving multiple pulses, such as estimating the date of Native American admixture in Rapa Nui \cite{moreno2014genome} or determining migration patterns in the Americas \cite{gravel2013reconstructing}. In these instances the expected decay of LD will become a mixture of exponentials. ROLLOFF and ALDER have limited resolution, as they can usually only infer the date of the most recent migration wave \cite{moorjani2011history}, or reject the hypothesis of a single pulse admixture \cite{loh2013inferring}.

ROLLOFF and ALDER use the information contained in pairs of sites by looking at the two-locus linkage disequilibrium between them. We use the information in triples of sites by considering three-locus LD. There are two ways of measuring the linkage between $n$ loci. Two-locus linkage disequilibrium decreases geometrically each generation as a result of recombination. \cite{bennett1952theory} defines $n$-locus linkage in a way that this property is maintained. Another property of two-locus LD is that it is equal to the covariance in the allele frequencies between the two sites. \cite{slatkin1972treating} defines $n$-locus LD analogously. For two and three loci, these two definitions coincide, but for four or more loci, they do not.

In this paper, we will use Bennett and Slatkin's definition of three-locus LD to look at the decay of weighted LD for three sites as a function of the genetic distance between them. We derive an equation that describes the decay of three-locus LD under an admixture history with multiple waves of migration. We then compare the results of coalescent simulations to this equation, and develop some guidelines for when admixture histories more complex than a single pulse can be resolved. Finally, we compute the our method for the Colombian and Mexican samples in the 1000 Genomes data set, using the Yoruba samples as a reference. Fitting a two-pulse model to data, we estimate admixture histories for the two populations which are qualitatively consistent with the results reported in \cite{gravel2013reconstructing}.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section{Model}
We use the same random union of gametes admixture model as in \cite{liang2014understanding}, which is itself an extension of the mechanistic admixture model formulated by \cite{verdu2011general}. In this model, two (or more) source populations contribute migrants to form an admixed population consisting of $2N$ haploid individuals. Each generation in the admixed population is formed through the recombination of randomly selected individuals from the previous generation, with some individuals potentially replaced by migrants from the source populations. For simplicity, we consider a model with only two source populations. Furthermore, the first source population only contributes migrants in the founding generation, $T$. The second source population contributes migrants in the founding generation and possibly in one or more generations thereafter. In generation $i$, for $i=T-1,\dots,0$ (before the present), a fraction $m_i$ of the admixed population is replaced by individuals from the second source population.

\section{Linkage Disequilibrium and Local Ancestry}
ROLLOFF and ALDER use the standard two-locus measure of LD between a SNP at positions $x$ and another SNP at position $y$, which is a genetic distance $d$ to the right, 
\begin{align}
D_2(d) = \text{cov}(H_x,H_y),
\label{D2}
\end{align}
where $H_x$ and $H_y$ represent the haplotype or genotypes of an admixed chromosome at positions $x$ and $y$. In the case of haplotype data, $H_{i,x}=1$ if the $i^\text{th}$ sample is carrying the derived allele at the SNP at position $x$, or is 0 otherwise. Alternatively, for genotype data, $H_{i,x}$ take on values from $\{0,1/2,1\}$ depending on the number of copies of the derived allele the $i^\text{th}$ sample is carrying at the SNP position $x$. 
We consider an additional site at position $z$, which is located a further genetic distance $d'$ to the right of $y$. The three-loci LD, as defined by as defined by \cite{bennett1952theory} and \cite{slatkin1972treating}, is given by
\begin{align}
	D_3(d,d')
	= \text{cov}(H_x,H_y,H_z)
	= \mathbb{E}[(H_x-\mathbb{E}H_x)(H_y-\mathbb{E}H_y)(H_z-\mathbb{E}H_z)].
\label{D3}
\end{align}
The LD in an admixed population depends on the genetic differentiation between the source populations and and its admixture history. Let $A_x$ represent the local ancestry at position $x$, with $A_x=1$ if $x$ is inherited from an ancestor in the first source population, and $A_x=0$ if $x$ is inherited from the second source population. We can compute the expectation of $D_3$ in terms of the three-point covariance function of $A_x$ and so separate out the effects of allele frequencies and local ancestry. We make the assumption that the alleles in the source populations are independent, so that  
\begin{align*}
	\text{cov}\left(H_x,H_y,H_z\right)
		&= \text{cov}\left(\mathbb{E}[H_x|A_x],\mathbb{E}[H_y|A_y],\mathbb{E}[H_z|A_z]\right).
\end{align*}
The background LD in unadmixed human populations decays to zero on a scale of tenths of centiMorgans, so this approximation is appropriate when $d$ and $d'$ are both larger than 0.5 cM. The conditional expectations above are the allele frequencies at each site in the admixed population, conditional on the local ancestry. These are given by $\mathbb{E}[H_x|A_x] = F_x+\delta A_x$, where $F_x$ is the allele frequency of locus $x$ in the first source population and $\delta_x$ is the difference of the allele frequencies of locus $x$ in the two source populations. Equation \ref{D3} becomes 
\begin{align}
	D_3(d,d')
		&= \text{cov}\left(f_x+\delta_xA_x,f_y+\delta_yA_y,f_z+\delta_zA_z\right)\nonumber \\	
		&= \delta_x\delta_y\delta_z \text{cov}(A_x,A_y,A_z).
	\label{local_ancestry}
\end{align}
A similar argument shows that $D_2(d)$ is proportional to the two-point covariance function of the local ancestry.

\subsection{Local Ancestry Covariance Functions}
If we take genetic drift into account, the three-point covariance function is random. To compute its expectation, we multiply out the covariance in equation \ref{D3} to get
\begin{align*}
	\mathbb{E}[\text{cov}(A_x,A_y,A_z)]
		&= \mathbb{E}[A_xA_yA_z]
		-\mathbb{E}[A_xA_y]\mathbb{E}[A_z]
		-\mathbb{E}[A_xA_z]\mathbb{E}[A_y]
		-\mathbb{E}[A_yA_z]\mathbb{E}[A_x]
		+2\mathbb{E}[A_x]\mathbb{E}[A_y]\mathbb{E}[A_z].
\end{align*}
Each one of these expectations on the right-hand side is the probability that one or more sites is inherited from an ancestor from first source population. We organize these products of probabilities in a column vector:
\begin{align*}
	\mathbf{v}_3 &= \left(\begin{array}{l}
	\mathbb{P}\{A_x=A_y=A_z=1\}\\ 
	\mathbb{P}\{A_y=A_z=0\}\mathbb{P}\{A_x=0\}\\
	\mathbb{P}\{A_x=A_z=0\}\mathbb{P}\{A_y=0\}\\
	\mathbb{P}\{A_x=A_y=0\}\mathbb{P}\{A_z=0\}\\ 
	\mathbb{P}\{A_x=0\}\mathbb{P}\{A_y=0\}\mathbb{P}\{A_z=0\}
	\end{array}\right),
\end{align*}
so that $\text{cov}(A_x,A_y,A_z) = (1,-1,-1,-1,2)\mathbf{v}_3$. There is one entry in $\mathbf{v}_3$ for each of the five ways in which the three markers at positions $x,y$, and $z$ can arranged on one or more chromosomes. In the founding generation $T$, this column vector is given by $\mathbf{v}_{3(T)} = (1-m_T,(1-m_T)^2,(1-m_T)^2,(1-m_T)^2,(1-m_T)^3)'$. The probabilities for subsequent generations can be found by left-multiplying drift, recombination, and migration matrices:
\begin{align*}
	\mathbf{v}_{3(i)} = \mathbf{D}_i \mathbf{L}\mathbf{U} \mathbf{v}_{3(i-1)},
\end{align*}
The matrices $\mathbf{D}_i$, $\mathbf{L}$, and $\mathbf{U}$ account for the effects of migration, drift, and recombination, respectively. The migration matrix is a diagonal matrix given by 
\begin{align*}
	\mathbf{D}_i &= \text{diag}(1-m_i,(1-m_i)^2,(1-m_i)^2,(1-m_i)^2,(1-m_i)^3).
\end{align*}
Its entries are the probabilities that one, two, or three chromosomes in the admixed population will not be replaced by chromosomes from the second source population in generation $i$. The lower triangular drift matrix 
\begin{align*}
	\mathbf{L}&=\frac{1}{4N^2}\left( \begin{array}{ccccc}
					 4N^2 		& 0 	& 0 & 0 & 0\\
					 2N 	& 2N-1 & 0 & 0 & 0\\
					 2N 	& 0 & 2N-1 & 0 & 0\\
					 2N 	& 0 & 0 & 2N-1 & 0\\
					 1 & 2N-1 & 2N-1 & 2N-1 & (2N-1)(2N-2)
				\end{array} \right)
\end{align*}
gives the standard Wright-Fisher drift transition probabilities between the states as a function of the population size $2N$. Finally, the upper triangular recombination matrix is determined by the recombination rates between the three sites:
\begin{align*}
	\mathbf{U} &=
		\left( \begin{array}{ccccc}
		e^{-d-d'} & (1-e^{-d})e^{d'} & (1-e^{-d})(1-e^{-d'}) & e^{-d}(1-e^{-d'}) & 0 \\
		0 & e^{-d'} & 0 & 0 & 1-e^{-d'} \\
		0 & 0 & 1-e^{-d}-e^{-d'}+2e^{-d-d'} & 0 & e^{-d}+e^{-d'}-2e^{-d-d'} \\ 
		0 & 0 & 0 & e^{-d} & 1-e^{-d} \\ 
		0 & 0 & 0 & 0 & 1 \end{array} \right)
\end{align*}
The covariance function is then given by
\begin{align}
	\mathbb{E}[\text{cov}(A_x,A_y,A_z)] &= \left(1,-1,-1,-1,2\right)\left(\prod_{i=0}^{T-1} \textbf{D}_i \mathbf{L}\mathbf{U}\right)\textbf{v}_{3(0)}.
	\label{cov}
\end{align} 
We can obtain an analogous equation for $\text{cov}(A_x,A_y)$, involving the migration, drift, and recombination matrices for two loci:
$$
	\mathbb{E}[\text{cov}(A_x,A_y)] = \left(1,-1\right)\left(\prod_{i=0}^{T-1} \textbf{D}_i \mathbf{L}\mathbf{U}\right)\textbf{v}_{2(0)}.
$$

In some cases, equation \ref{cov} simplifies further. In a one-pulse migration model, in which $m_T=M$ and is there after 0, the $\mathbf{D}_i$'s become identity matrices, and we get the closed from expression 
$$
	\mathbb{E}[\text{cov}(A_x,A_y,A_z)] = M(1-M)(1-2M)\left(1-\frac{1}{2N}\right)^T\left(1-\frac{2}{2N}\right)^Te^{-T(d+d')}.
$$
This is because $(1,-1,-1,-1,2)$ is a left eigenvector of both $\mathbf{L}$ and $\mathbf{U}$, with corresponding eigenvectors $(1-1/2N)(1-2/2N)$ and $\exp(-d-d')$. Note that when $M=0$, the covariance function will be identically 0. Another case is a two pulse model in which we ignore the effects of genetic drift. In this model, admixture only occurs $T$ and $T_2$ generations before the present, so that $m_T=M_1,m_{T'}=M_2$, and all other $m_i$'s are 0. Making the substitution $T_1=T-T_2$, the right hand side of equation \ref{cov} becomes 
\begin{align}
		&(1-M_1)(1-M_2)e^{-T_2(d+d')}\left[M_2(1-M_1)^2-2M_2^2(1-M_1)^2+M_1(1-2M_1)e^{-T_1(d+d')}\right.\nonumber\\
		&\ \ \left.-M_1M_2(1-M_1)\left(e^{-M_1d}+e^{-M_1 d'}+\left(1-e^{-d}-e^{-d'}+2e^{-d-d'}\right)^{T_1}\right) \right].
\label {ash_2p}
\end{align} 
The corresponding expression for the two-point covariance function is given by 
\begin{align}
		(1-M_1)(1-M_2)e^{-T_2d}\left(M_2-M_1M_2+M_1e^{-T_1d}\right),
\label{alder_2p}
\end{align} 
which is a mixture of two exponentials. The relative complexity of equation \ref{ash_2p} is actually a feature, as it makes detecting the presence of the second pulse of admixture easier.
\section{Weighted Linkage Disequilibrium}
As \cite{loh2013inferring} noted, we cannot use the LD in the admixed population directly, because the allele frequency differences in the source populations can be of either sign. However, if we compute expectation of the product of the LD with the product of the allele frequency differences, using equation \ref{local_ancestry} we obtain
$$
\mathbb{E}\left[\delta_x\delta_y\delta_z D_3(d,d')\right]=\mathbb{E}[\delta_x^2\delta_y^2\delta_z^2]\mathbb{E}[\text{cov}(A_x,A_y,A_z)],
$$
because the local ancestry in the admixed sample is independent of the allele frequencies in the admixed population. This expectation of the weighted LD is non-zero, and can be estimated by aggregating over triples of SNPs which are separated by distances of approximately $d$ and $d'$. The LD term can be estimated from the admixed population, while the $\delta$'s can be estimated from reference populations which are closely related to the two source populations.

We arrange the data from the admixed samples in an $n\times S_n $ matrix $\mathbf{H}$, where $n$ is the number of admixed haplotypes/genotypes, and $S_n$ is the number of segregating sites in the sample. For ease of notation, we assume that the positions are given in units which make the unit interval equal to the desired bin resolution.

For a given $d$ and $d'$ the set of SNP triples we use in the estimator for the weighted LD is 
$$
	S[d,d'] = \left\{x,y,z: d\leq  x-y < d+1 \text{ and } d' \leq y-z < d'+1 \right\}.
$$
Let $w_x$ be the difference in the empirical allele frequencies in two reference populations and let $f_x$ be empirical allele frequency in the admixed population. An unbiased estimator of the weighted LD is
\begin{align*}
	\hat{a}[d,d']&=
		\frac{1}{|S[d,d']|}
		\sum_{x,y,z\in S[d,d']}\frac{n\sum_{i=1}^n
			w_a w_b w_c(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)}{(n-1)(n-2)}.
\end{align*}

\section{Algorithm}
Directly computing $\hat{a}[d,d']$ over the set $d,d'\in \{0,1,\dots,P\}^2$ would be cubic in the number of segregating sites, but as is the case with ALDER, we can use using a fast Fourier transform (FFT) to approximate $\hat{a}$, giving an algorithm whose run-time is instead linear in the number of segregating sites. We first rearrange $\hat{a}$ to get
\begin{align*}
	\hat{a}[d,d']
		&=\frac{n}{(n-1)(n-2)}
		\frac{\sum_{i=1}^n\sum_{x,y,z\in S[d,d']}
			\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)}
		{\sum_{x,y,z\in S[d,d']}1},
\end{align*}
and define sequences $b_i[d]$ and $c[d]$ by binning the data and then doubling the length by padding with $P$ zeros,
\begin{align*}
	b_i[d] &= \left\{
     \begin{array}{ll}
       \sum_{x: d\leq \lfloor{x}\rfloor < d+1}\delta_x(H_{i,x}-f_x) & : 0\leq d < P\\
       0 & : P\leq d <2P
     \end{array}
   \right.\\
  	c[d] &= \left\{
     \begin{array}{ll}
       |\{x: d\leq \lfloor{x}\rfloor < d+1\}| & : 0\leq d < P\\
       0 & : P\leq d <2P
     \end{array}
   \right.
\end{align*}
We can approximate $|S[d,d']|$ and the $n$ sums in the numerator of $\hat{a}[d,d']$ in terms of convolutions of these sequences:
$$
	|S[d,d']| \approx \sum_{w=0}^P c[w]c[w+d]c[w+d+d']
$$
$$
	\sum_{x,y,z\in S[d,d']}\delta_x\delta_y\delta_z(H_{i,x}-f_x)(H_{i,y}-f_y)(H_{i,z}-f_z)
		\approx
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'].
$$
These convolutions can be efficiently computed with an FFT, since under a two-dimensional discrete Fourier transform, 
$$
	\sum_{w=0}^P b_i[w]b_i[w+d]b_i[w+d+d'] \leftrightarrow B_i[j]\bar{B_i}[k]B_i[k-j],
$$
where $B_i$ is the (one-dimensional) discrete Fourier transform of $b$, and $B_i[-j]$ is the $j^\text{th}$ to last most element of $B_i$. Summing over $i$ and taking the inverse discrete Fourier transform, we can approximate the discrete Fourier transform of numerator of  $\hat{a}$. We use the same method applied to $c$ to approximate the denominator of $\hat{a}$. Because the number of bins is generally much less than the number of segregating sites, the rate-limiting step of this algorithm is the binning step to form $c$ and the $b_i$'s, which is $O(S_n)$, rather than the FFTs, which are $O(P^2\log(P))$.

When using only the admixed population itself as a reference population, the method described above will be biased if the same samples are used to estimate both the linkage disequilibria and the weights. We cannot efficiently compute a polyache statistics like \cite{loh2013inferring}. At the cost of some power, we instead adopt the approach of \cite{pickrell2012inference} and separate the admixed population into two equal-sized groups. We then use one group to estimate the weights, and the other group to estimate linkage disequilibrium, and vice versa. This gives gives two unbiased estimates for the numerator of $\hat{a}$, which we then average.
\subsection{Fitting the Two-Pulse Model}
We fit equation \ref{alder_2p} to the estimates of the weighted LD using non-linear least squares, with two modifications. We added a proportionality constant to account for the expected square allele frequency difference between the source populations. We also subtracted out an affine term in the weighted LD which is due to population substructure \cite{loh2013inferring}. We estimated this by computing the three-way covariance between triples of chromosomes. We use the jackknife to obtain confidence intervals for the resulting estimates by leaving out each chromosome in turn and refitting on the data for the remaining chromosomes.
\section{Simulations}
We used the program macs \cite{chen2009fast} to generate two source populations which diverged 4000 generations ago and a coalescent simulation to generate an admixed population from the two source populations according to two-pulse and constant admixture models. We sampled 50 diploid individuals from the admixed and two source populations, each consisting of 20 chromosomes of length 1 Morgan. The effective population size was $2N=1000$ for the admixed population and two source populations. Using a two pulse model, we varied the migration probabilities and timings for each pulse to examine the accuracy of equation \ref{alder_2p}. We also simulated data for a model with a constant rate of admixture each generation, and compared this to the predictions made by equation \ref{cov}.
\section{Data Set}
We computed the weighted LD for the Mexican and Columbia populations in the first phase of the 1000 Genomes data set. These consisted of 66 individuals from Los Angeles and 60 individuals from Medellin, respectively. We used the 88 Yoruba samples as the one reference population. We computed the weighted LD on the genotypes to avoid effects of phasing errors.
% Results and Discussion can be combined.
\section{Discussion}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
\subsection{Simulations}

\begin{figure}
\includegraphics[scale=.6]{Ts.eps}\caption{
{\bf Predicted weighted LD surfaces from simulations and theory for varying admixture times.} The heat maps are from simulations and the contours are plotted from equation \ref{D3}. The two admixture probabilities were fixed at $m_1=m_2=.2$ and the the times of the two admixture pulses, $T_1$ and $T_2$, were varied. Each square covers the range $0.5 \text{ cM }<d,d'<20\text{ cM}$. When time of the more recent pulse is greater than half of that of the more ancient pulse, i.e. $2T_1>T_1+T_2$, the contours of the resulting weighted LD surface are straight, making it difficult to distinguish from the weighted LD surface produced by a one-pulse admixture scenario.
}
\label{Ts}
\end{figure}

\begin{figure}
\includegraphics[scale=.6]{ms.eps}
\caption{
{\bf Predicted weighted LD surfaces from simulations and theory.} The heat maps are from simulations and the contours are plotted from equation \ref{D3}. The two admixture times were fixed at 2 and 12 generations ago ($T_1=10$ and $T_2=2$) while the admixture probabilities were varied. Each square covers the range $0.5 \text{ cM }<d,d'<20\text{ cM}$. As the total admixture proportion $m_2+m_1(1-m_2)$ increases above 0.5, the concavity of the contours flips. Weighted LD surfaces for $m_1>0.5$ or $m_2>0.5$ are not shown, but are qualitatively similar to the surfaces on the lower and rightmost sides.  
}
\label{ms}
\end{figure}

\begin{figure}
\includegraphics[scale=.8]{continuous.eps}
\caption{
{\bf Weighted LD surfaces produced by constant admixture.} The heat maps are from simulations and the contours are from equation XX. In all six plots, admixture stopped 5 generations before the present. Each square covers the range $0.5 \text{ cM }<d,d'<20\text{ cM}$. We varied the time of the beginning of the admixture and the total admixture probability. The admixture probability for each generation was constant, and chosen so that the total admixture proportion was either $0.3$ or $0.7$. When the admixture is spread over 5 generations (the leftmost column), the resulting weighted LD surface is similar to a one-pulse weighted LD surface. For longer durations, the weighted LD surfaces are similar to those produced by two pulses of admixture.
}
\label{continuous}
\end{figure}

We find there is a generally a close match between our equations and the simulated data under both under two pulse admixture scenarios (figures \ref{Ts} and \ref{ms}) and constant admixture scenarios (figure \ref{continuous}). The exception is when the total admixture proportion $M_2+M_1(1-M_2)$ is close to 0.5. As the total admixture proportion increases above 0.5, the contours for equation \ref{D3} flip from being concave down to concave up. This transition can be seen by comparing the upper left side of figure \ref{ms} to its lower right. At this threshold, the contours of the estimated weighted LD depend on the actual admixture fractions of the samples, which may differ from the expectation as a result of genetic drift. This mismatch between theory and simulations is most evident in figure \ref{ms}, for $m_1=0.1,m_2=0.4$ and $m_1=0.2,m_2=0.4$.

When there is continuous admixture scenario, the shape of the weighted LD surface depends on both the duration and total amount of admixture. When the duration is short, the weighted LD surfaces are indistinguishable from the weighted LD surfaces produced by one pulse of migration. As the duration increases, the contours of the weighted LD surface become more curved. The contours are concave up when the total proportion is greater than $50\%$ and concave down when it is less. When the total proportion is exactly $50\%$, the amplitude of the weighted LD surface is much smaller than the sampling error.

For two pulse models, the effects of the second pulse of migration only become evident when temporal spacing between the pulses is large enough ($T1 > T2$). Otherwise, the resulting weighted LD surface cannot be distinguished from the weighted LD surface produced by one pulse of admixture. As in the case of continuous admixture the concavity of the surface contours is determined by the total admixture proportion.

\begin{figure}
\includegraphics[scale=.6]{estimates.eps}
\caption{
{\bf Accuracy of estimates of $T_1$ as a function of other parameters.} Nine admixture scenarios, $T_1\in \{5,10,20\}$ and $T_2\in\{2,5,10\}$, were simulated 100 times each. The admixture probabilities were fixed at $M_1=0.3$ and $M_2=0.2$. The colored bars give the medians of estimates for each of these nine cases, the boxes delimit the interquartile range, and the whiskers extend out to 1.5 times the interquartile range. As the time between the two pulses of admixture increases, the error in the estimates decreases. Consistent with the simulations shown in figure \ref{Ts}, there is limited power to estimate the time of the more ancient admixture pulse when $T_2>T_1$.
}
\label{estimates}
\end{figure}

These qualitative about the similarity between one pulse and two pulse admixture scenarios are borne out by simulations of the estimation error, shown in figure \ref{estimates}. When the spacing between the two pulses is small relative to their age, the median of the estimates of the timing of the second pulse is close to the true value, but the interquartile range is large. Moreover, the best fit often lies on a boundary of the parameter space which is equivalent to a one pulse admixture model. When the spacing between the pulses is larger, the estimates for the timing of the older pulse before more precise.
\subsection{1000 Genomes}
\begin{figure}
\includegraphics[scale=.6]{MXL.eps}
\caption{
{\bf Weighted LD surface for Mexican samples with Yoruba as reference.} The model with the best fit is two pulses from the non-Yoruba source population at $T_1+T_2=12.3 \pm 3.3$ and $T_2=9.9\pm 2.7$ generations ago. The jackknife confidence intervals for the times of these two pulses overlap. 
}
\label{ASH_MXL}
\end{figure}

\cite{gravel2013reconstructing} have previously analyzed the 1000 Genomes data that we computed weighted LD surfaces for. For the Mexican samples, they found a small but consistent amount of African ancestry, which appeared in the population 15 generations ago, with continuing contributions from European and Native American populations since that date, but no African migration. In fitting a two-pulse model to the Mexican weighted LD surface (figure \ref{ASH_MXL}), we estimated that the two pulses occurred $12.3\pm3.3$ and $9.9\pm2.7$ generations ago. These confidence intervals overlap, and so we cannot reject a one-pulse admixture history. This is not quite consistent with the constant migration model that \cite{gravel2013reconstructing} found, but as we have seen from simulations, it is hard to distinguish a constant migration model from a one-pulse model when the duration of the migration is short.

\begin{figure}
\includegraphics[scale=.6]{CLM.eps}
\caption{
{\bf Weighted LD surface for Colombian samples with Yoruba as reference} The two-pulse model that fits best is two pulses of non-Yoruba admixture at $T_1+T_2=11.8\pm 1.2$ and $T_2=2.64 \pm 0.08$ generations ago. The jackknife confidence intervals for the times of these two pulses do not overlap. The amplitude of this weighted LD surface is approximately ten times larger than that of the Mexican samples. This a result of larger proportion of Yoruba ancestry in the Colombian samples.
}
\label{ASH_CLM}
\end{figure}

The weighted LD surface for the Columbia samples is shown in figure \ref{ASH_CLM}. From this, we estimated two pulses of non-Yoruba migration at $11.8\pm 1.2$ and $2.64\pm0.08$ generations before the present. \cite{gravel2013reconstructing} also inferred two pulses of admixture, corresponding to 3 and 9 generations ago. The weighted LD surface of the Colombian samples has contours which are strongly concave up, in contrast to those of the Mexican samples. 
\subsection{Comparison to Existing Methods}
Compared to existing weighted LD methods, our our method uses more information in the data because it compares triples of SNPs instead of pairs. This gives our method the ability to infer admixture histories more complex than a one-pulse model. However, this comes at the price of greater estimation variances. ALDER and ROLLOFF can make estimates from just tens of samples, while our method requires hundreds of samples. Part of this difference can be attributed to the fact that ALDER and ROLLOFF make inferences over a smaller class of models, but the main reason arises from the fact that the existing two models are estimating second moments of the data, while we are estimating third moments. The variance of these estimates are both inversely proportional to the sample size, but the constants for estimating third moments are larger. As data becomes more readily available, this disadvantage should disappear.

\bibliographystyle{plos_template}
\bibliography{}

\begin{figure}
\includegraphics[scale=.6]{alder_Ts.eps}
\caption{
{\bf Extra figure} Corresponding alder curves for two-pulse admixture with varying pulse times. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Red lines are $T_2=2$, Green lines $T_2=5$, and blue lines are $T_2=10$.
}
\end{figure}

\begin{figure}
\includegraphics[scale=.6]{alder_ms.eps}
\caption{
{\bf Extra figure} Corresponding alder curves for two-pulse admixture with varying migration probabilities. Morgans on $x$-axis and $\log$ Alder scores on $y$-axis. Increasing red values are higher $M_1$'s and increasing green values are higher $M_2$'s  (I will probably need to choose a better color scheme for this).
}
\end{figure}

\end{document}